Stateless Model Checking with Data-Race Preemption Points

### 2015-11-03 pldi-16/di.tex
Implement jiri's comments

Iterative Deepening depends heavily on state-space estimation \cite{estimation}
to understand which state spaces are likely to complete on time,
in advance of actually testing each interleaving within.
The purpose is to make decisions automatically about when to defer exploration of a state space,
so an inexpert user need only provide their total CPU budget as a test parameter,
and to enable completing appropriately-sized state spaces within that budget.

@@

For the rest of the paper, we abbreviate {\em preemption point} (PP),
%For the remainder of the paper, we will abbreviate {\em preemption point} (PP),
{\em model checking} (MC),
{\em single-state-space MC} (SSS-MC) (i.e., the approach of prior work),
Dynamic Partial Order Reduction (DPOR), and {\em state space estimate} (ETA).

Please note the difference between data-race {\em candidates} and data-race {\em bugs}.
Because data-race analysis is false-positive-prone,
we classify unprotected access pairs separately from concrete observable failures, % "concrete"?
calling such pairs {\em data race candidates}.
Should a future interleaving, preempting during those accesses,
lead to a failure (e.g. assertion or deadlock), then we report a {\em data-race bug}.
Otherwise, if the access pair can be reordered, but does not produce a failure under any interleaving, it is a {\em benign data race}.
If they cannot be reordered at all, it is a {\em false positive}.
%TODO: Don't do this, below. Scan paper for uses of 'false positive' where you mean to say 'benign' or 'both'.
%For brevity, we count benign data races as a subset of false positives.

@@

\subsection{Verifying data races as buggy or benign} % FIXME: dont say 'benign'; also, do you have enough left to say that this can even still be a full subsection?

@@

% TODO: put colour in this code
\newcommand\PendingJobs{\ensuremath{\mathcal{P}}}
\newcommand\SuspendedJobs{\ensuremath{\mathcal{S}}}
\newcommand\GetETA[1]{ETA(#1)}
\newcommand\GetPPSet[1]{PPSet(#1)}
\begin{algorithm}[t]
	\SetKwInOut{Input}{Input}
	%\textbf{Function} GetBestJob($j_0$, PendingJobs, SuspendedJobs): \\
	\Input{$j_0$, the currently-running job}
	%\Input{$eta$, $j_0$'s predicted completion time}
	\Input{\PendingJobs, the list of pending jobs, sorted decreasingly by heuristic priority}
	\Input{\SuspendedJobs, the list of already-suspended jobs, sorted increasingly by ETA}
	\If{\GetETA{$j_0$} $<$ HeuristicETAFactor $\times$ TimeLeft()}{
		// common case: job will probably finish on time \\
		return $j_0$
	}
	\ForEach{job $j_P \in$ \PendingJobs}{
		// don't run a pending job if a subset of it is already suspended; it would find the same fate! \\
		\If {$\forall j_S \in$ \SuspendedJobs, \GetPPSet{$j_S$} $\not\subset$ \GetPPSet{$j_P$}}{
			return $j_P$
		}
	}
	%// no pending jobs; maybe resume a suspended job \\
	\ForEach{job $j_S \in$ \SuspendedJobs}{
		\If{\GetPPSet{$j_0$} $\not\subset$ \GetPPSet{$j_S$}
			$\land$
			\GetETA{$j_0$} $>$ \GetETA{$j_S$}}{
			// $j_S$ is tempting, but if a subset of it is also suspended, don't run the larger one first \\
			\If{$\forall j_{S2} \in$ \SuspendedJobs, \GetPPSet{$j_{S2}$} $\not\subset$ \GetPPSet{$j_S$}}{
				return $j_S$
			}
		}
	}
	// $j_0$'s ETA was bad, but nothing else was better \\
	return $j_0$
	\caption{Suspending exploration of a too-large state space in favour of a potentially smaller one.}
	\label{alg:shouldworkblock}
\end{algorithm}

With a limited CPU budget, we must avoid running tests likely to time out.
When a MC instance reports an ETA too high,
we compare other yet-untested PP sets to find another job more likely to complete in time.
Our method for doing so, listed in Algorithm~\ref{alg:shouldworkblock}, is the heart of Iterative Deepening\footnote{
Though its worst-case performance is $O(mn)$ in the
%number of pending and suspended jobs,
sizes of $\mathcal{P}$ and $\mathcal{S}$,
in practice the non-constant portion beyond line 4 runs very infrequently
and is negligible compared to the exponentially-sized state spaces.}.
Its main feature is understanding that if \GetPPSet{$j_1$} $\subset$ \GetPPSet{$j_2$},
and $j_1$ is suspended,
then $j_2$'s state space is guaranteed to be strictly larger, so $j_2$ will take at least as long.
Hence we should avoid testing $j_2$ unless $j_1$'s ETA improves after further time spent on it.
%reveals that it might finish in time after all.
Similarly, when some job finds a bug, we cancel all pending superset jobs, as they would find the same bug.

%
We also account for the inherent inaccuracy of ETA estimates.
Line 1 heuristically scales up the time remaining to avoid suspending jobs too aggressively
in case their ETAs are actually overestimates.
Lines 12-14 account for the
%bizarre
possibility that among two suspended jobs,
%given two jobs,
%%$j_1,j_2$,
\GetPPSet{$j_1$} $\subset$ \GetPPSet{$j_2$}
but
\GetETA{$j_1$} $>$ \GetETA{$j_2$}.
This can arise because estimates tend to get more accurate over time,
and $j_1$ perhaps ran much longer before suspending.
% In such scenarios,
We heuristically assume the smaller job's ETA is more accurate
to avoid repeatedly resuming larger jobs briefly while their ETAs only become worse
(it lets us avoid thrashing in \quicksand),

@@

We chose \landslide~\cite{landslide} as our stateless model checker due to its ability to trace program execution at the granularity of individual instructions and memory accesses, which dynamic data-race detection requires.
\landslide~features DPOR,
%\cite{dpor},
state space estimation, and a hybrid lockset/happens-before data-race analysis.
% TODO: \cite{fairstatelessmc} instead or in addition?
It avoids state space cycles (e.g. ad-hoc synchronization with {\tt yield()} or even {\tt xchg} loops) with a heuristic similar to Fair-Bounded Search \cite{bpor}. % TODO: devote a paragraph to this, after talking about drs in locks
% this line can be cut if space is needed
It can test both user- and kernel-level code, although is limited to timer-driven nondeterminism.
Its bug-detection metrics include assertion failure, deadlock, segfault, heap checking (like Valgrind~\cite{valgrind}), and a heuristic infinite loop/livelock check.

{\bf Restricting PPs with stack trace predicates.}
Most MCs preempt indiscriminately on any sync API call, regardless of the call-site.
However, when testing a particular module in a large codebase,
the user is likely uninterested in PPs arising from other modules.
\landslide~provides the {\tt within\_function} configuration command for a user to identify which call-sites matter most.
Before inserting a PP, \landslide~requires at least one argument to {\tt within\_function} to appear in the current thread's stack trace.
The {\tt without\_function} directive works similarly, but as a blacklist.
Multiple invocations can be used; later ones take precedence.
%\cite{landslide} provides further detail on this feature.

{\bf Data races in lock implementations.}
Prior work data race tools recognize the implementations of sync primitives to avoid spuriously flagging memory accesses resulting from the lock implementation itself \cite{tsan}.
Assuming the locks are already correct enables productive data-race analysis on the rest of the codebase.
With testing limited to one execution, even if you wished to test for lock bugs, data-race analysis would still uselessly flag every single access pair in the lock implementation, requiring human attention to verify.
However, Iterative Deepening can automatically verify a large quantity of data-race candidates as benign.
Hence, we extended \landslide~with a custom option to change the lock-set tracking to include accesses from {\tt mutex\_lock()} and {\tt mutex\_unlock()} in the analysis. (Accesses from other sync functions, such as {\tt cond\_wait()}, would either be included already, or be protected by an internal mutex.)
% TODO: Add a code sample showing how the lockset boundaries change? If there's room?

@@

{\bf Communication protocol.}
The interface to \landslide~, which any similar MC could implement, is in two parts.
First, when starting each job, it writes a configuration file declaring which PPs to use,
% can lose this line due to space
among other options such as mutex-testing mode,
passed as an argument to \landslide.
Then, a dedicated \quicksand~thread communicates with the \landslide~process via message-passing on a FIFO pipe.
\landslide~sends a message after each new interleaving is tested to report updated progress and ETA,
whenever a new data-race candidate is found, and whenever a bug is found.
\quicksand~in turn replies whether the test should suspend/resume due to too high ETA, or quit due to timeout.

@@

We classify data-race candidates as {\em single-order} or {\em both-order} \cite{portend}
based on whether \landslide~observed the access pair ordered one or both ways while testing the original state space,
Single-order candidates are much more likely to be false-positives
(though we cannot say for sure before preempting during the access itself, hence the heuristic),
so we prioritize jobs with both-order data-race PPs more highly.
For single-order races, we do not add PPs for the later access at all;
should they be needed, preempting on the first access will suffice to upgrade it to both-order.

@@

When \quicksand~receives a data race report, it adds two new jobs to its workqueue:
a ``small'' job to preempt on the racing instruction only,
and a ``big'' job to preempt on that instruction as well as each PP used by the reporting job.
%
Hence each {\em pair} of racing accesses will spawn four new jobs, as shown in Figure~\ref{fig:new-dr-jobs}.
Rather than adding a maximal job with both new PPs at once, we prefer to add smaller jobs which have a higher chance of completing in time.
If those state spaces are bug-free, they will in turn add the maximal job later.
%
The rationale of spawning multiple jobs is that which will be more fruitful cannot be known in advance:
while the big job risks not completing in time,
the small job risks missing the data race entirely if the original PPs were required to expose it.
In practice, we observed some bugs found quickly by these small jobs, and other bugs missed by the small jobs found eventually by the big jobs,
which motivates the need for Iterative Deepening to decide at runtime which to prioritize.
% TODO: Put numbers here.

>>>>

Iterative Deepening depends heavily on state-space estimation \cite{estimation}
to understand which state spaces are likely to complete on time,
in advance of actually testing each interleaving within.
The purpose is to make decisions automatically about when to defer exploration of a state space,
so an inexpert user must provide only their total CPU budget as a test parameter,
and to enable completing appropriately-sized state spaces within that budget.

@@

For the rest of the paper, we abbreviate {\em preemption point} (PP),
%For the remainder of the paper, we will abbreviate {\em preemption point} (PP),
{\em model checking} (MC),
{\em single-state-space model checking} (SSS-MC),% (i.e., the approach of prior work),
Dynamic Partial Order Reduction (DPOR), and {\em state space estimate} (ETA).

SSS-MC indicates the approach of prior model checking tools:
the set of PPs is fixed in advance, and the tool commits to testing every interleaving available via those PPs.
%Reduction techniques are often used to skip equivalent interleavings \cite{dpor}, and search-ordering strategies
Many techniques exist to skip equivalent interleavings and order the search to uncover bugs faster \cite{dpor,demeter,chess-icb,gambit},
but dynamic analysis cannot add new PPs or remove ineffective ones.

Please note the difference between data-race {\em candidates} and data-race {\em bugs}.
Because data-race analysis is prone to false-positives,
we classify unprotected access pairs separately from concrete observable failures, % "concrete"?
calling such pairs {\em data race candidates}.
Should a future interleaving, preempting during those accesses,
lead to a failure (e.g. assertion or deadlock), then we report a {\em data-race bug}.
Otherwise, if the access pair can be reordered, but does not produce a failure under any interleaving, it is a {\em benign data race}.
If they cannot be reordered at all, it is a {\em false positive}.
%TODO: Don't do this, below. Scan paper for uses of 'false positive' where you mean to say 'benign' or 'both'.
%For brevity, we count benign data races as a subset of false positives.

@@

\subsection{Classifying data races as buggy or benign} % FIXME: dont say 'benign'; also, do you have enough left to say that this can even still be a full subsection?

@@

% TODO: put colour in this code
\newcommand\PendingJobs{\ensuremath{\mathcal{P}}}
\newcommand\SuspendedJobs{\ensuremath{\mathcal{S}}}
\newcommand\GetETA[1]{ETA(#1)}
\newcommand\GetPPSet[1]{PPSet(#1)}
\begin{algorithm}[t]
	\SetKwInOut{Input}{Input}
	%\textbf{Function} GetBestJob($j_0$, PendingJobs, SuspendedJobs): \\
	\Input{$j_0$, the currently-running job}
	%\Input{$eta$, $j_0$'s predicted completion time}
	\Input{\PendingJobs, the list of pending jobs, sorted decreasingly by heuristic priority}
	\Input{\SuspendedJobs, the list of already-suspended jobs, sorted increasingly by ETA}
	\If{\GetETA{$j_0$} $<$ HeuristicETAFactor $\times$ TimeLeft()}{
		// Common case: Job is expected to complete. \\
		return $j_0$
	}
	\ForEach{job $j_P \in$ \PendingJobs}{
		// Don't run a pending job if a subset of it is already suspended; its ETA would be at least as bad. \\
		\If {$\forall j_S \in$ \SuspendedJobs, \GetPPSet{$j_S$} $\not\subset$ \GetPPSet{$j_P$}}{
			return $j_P$
		}
	}
	%// no pending jobs; maybe resume a suspended job \\
	\ForEach{job $j_S \in$ \SuspendedJobs}{
		\If{\GetPPSet{$j_0$} $\not\subset$ \GetPPSet{$j_S$}
			$\land$
			\GetETA{$j_0$} $>$ \GetETA{$j_S$}}{
			// If a subset of $j_S$ is also suspended, don't run the larger one first. \\
			\If{$\forall j_{S2} \in$ \SuspendedJobs, \GetPPSet{$j_{S2}$} $\not\subset$ \GetPPSet{$j_S$}}{
				return $j_S$
			}
		}
	}
	// $j_0$'s ETA was bad, but nothing else was better. \\
	return $j_0$
	\caption{Suspending exploration of a state space in favour of a potentially smaller one.}
	\label{alg:shouldworkblock}
\end{algorithm}

With a limited CPU budget, we must avoid running tests that are likely to time out.
When a MC instance reports an ETA too high,
we compare other yet-untested PP sets to find another job more likely to complete in time.
Our method for doing so, listed in Algorithm~\ref{alg:shouldworkblock}, is the heart of Iterative Deepening\footnote{
Though its worst-case performance is $O(mn)$ in the
%number of pending and suspended jobs,
sizes of $\mathcal{P}$ and $\mathcal{S}$,
in practice the non-constant portion beyond line 4 runs very infrequently
and is negligible compared to the exponentially-sized state spaces.}.
Its main feature is understanding that if \GetPPSet{$j_1$} $\subset$ \GetPPSet{$j_2$},
and $j_1$ is suspended,
then $j_2$'s state space is guaranteed to be strictly larger, so $j_2$ will take at least as long.
Hence we should avoid testing $j_2$ unless $j_1$'s ETA improves over time.
%reveals that it might finish in time after all.
Similarly, when some job finds a bug, we cancel all pending superset jobs, as they would find the same bug.

%
We also account for the inherent inaccuracy of ETA estimates.
Line 1 heuristically scales up the time remaining to avoid suspending jobs too aggressively
in case their ETAs are actually overestimated.
Lines 12-14 account for the
%bizarre
possibility that among two suspended jobs,
%given two jobs,
%%$j_1,j_2$,
\GetPPSet{$j_1$} $\subset$ \GetPPSet{$j_2$}
but
\GetETA{$j_1$} $>$ \GetETA{$j_2$}.
This can arise because estimates tend to get more accurate over time,
and $j_1$ perhaps ran much longer before suspending.
% In such scenarios,
We heuristically assume the smaller job's ETA is more accurate
to avoid repeatedly resuming larger jobs briefly while their ETAs only become worse
(it lets us avoid thrashing in \quicksand).

@@

We chose \landslide~\cite{landslide} as our stateless model checker due to its ability to trace program execution at the granularity of individual instructions and memory accesses, which dynamic data-race detection requires.
\landslide~implements DPOR,
%\cite{dpor},
state space estimation, and a hybrid lockset/happens-before data-race analysis.
% TODO: \cite{fairstatelessmc} instead or in addition?
It avoids state space cycles (e.g. ad-hoc synchronization with {\tt yield()} or even {\tt xchg} loops) with a heuristic similar to Fair-Bounded Search \cite{bpor}. % TODO: devote a paragraph to this, after talking about drs in locks
% this line can be cut if space is needed
It can test both user-level and kernel-level code, although is limited to timer-driven nondeterminism.
Its bug-detection metrics include assertion failure, deadlock, segfault, heap checking (like Valgrind~\cite{valgrind}), and a heuristic infinite loop/livelock check.

{\bf Restricting PPs with stack trace predicates.}
Most MCs preempt indiscriminately on any sync API call, regardless of the call-site.
However, when testing a particular module in a large codebase,
the user is likely uninterested in PPs arising from other modules.
\landslide~provides the {\tt within\_function} configuration command for a user to identify which call-sites matter most.
Before inserting a PP, \landslide~requires at least one argument to {\tt within\_function} to appear in the current thread's stack trace.
%The {\tt without\_function} directive works similarly, but as a blacklist.
The {\tt without\_function} directive is the dual of {\tt within\_function}, indicating a blacklist.
Multiple invocations can be used; later ones take precedence.
%\cite{landslide} provides further detail on this feature.

{\bf Data races in lock implementations.}
Data race tools in prior work \cite{tsan,portend} recognize the implementations of sync primitives to avoid spuriously flagging memory accesses resulting from the lock implementation itself.
Assuming the locks are already correct enables productive data-race analysis on the rest of the codebase.
With testing limited to one execution, even if you wished to test for lock bugs, data-race analysis would still uselessly flag every single access pair in the lock implementation, requiring human attention to verify.
However, Iterative Deepening can automatically verify a large quantity of data-race candidates as benign.
Hence, we extended \landslide~with a custom option to change the lock-set tracking to include accesses from {\tt mutex\_lock()} and {\tt mutex\_unlock()} in the analysis. (Accesses from other sync functions, such as {\tt cond\_wait()}, would either be included already, or be protected by an internal mutex.)
% TODO: Add a code sample showing how the lockset boundaries change? If there's room?

@@

{\bf Communication protocol.}
The interface to \landslide~, which any similar MC could implement, has two parts.
First, when starting each job, \quicksand~creates a configuration file declaring which PPs to use,
% can lose this line due to space
among other options such as mutex-testing mode,
passed as an argument to \landslide.
Then, a dedicated \quicksand~thread communicates with the \landslide~process via message-passing.%on a FIFO pipe.
\landslide~messages \quicksand~after testing each new interleaving to report updated progress and ETA,
whenever a new data-race candidate is found, and whenever a bug is found.
\quicksand~in turn replies whether the test should suspend/resume due to too high ETA, or quit due to timeout.

@@

We classify data-race candidates as {\em single-order} or {\em both-order} \cite{portend}
based on whether \landslide~observed the access pair ordered one or both ways while testing the original state space,
Single-order candidates are much more likely to be false-positives
(though we cannot say for sure before preempting during the access itself, hence the heuristic),
so we prioritize jobs with both-order data-race PPs.
For single-order races, we do not add PPs for the later access at all;
should they be needed, preempting on the first access will suffice to upgrade it to both-order.

@@

When \quicksand~receives a data race report, it adds two new jobs to its workqueue:
a ``small'' job to preempt on the racing instruction only,
and a ``big'' job to preempt on that instruction as well as each PP used by the reporting job.
%
Hence each {\em pair} of racing accesses will spawn four new jobs, as shown in Figure~\ref{fig:new-dr-jobs}.
Rather than adding a maximal job with both new PPs at once, we prefer to add smaller jobs which have a higher chance of completing in time.
If those state spaces are bug-free, they will in turn add the maximal job later.
%
The rationale of spawning multiple jobs is that which will be more fruitful cannot be known in advance:
while the big job risks not completing in time,
the small job risks missing the data race entirely if the original PPs were required to expose it.
In practice, we observed some bugs found quickly by these small jobs, and other bugs missed by the small jobs found eventually by the big jobs,
which motivates the need for Iterative Deepening to decide at runtime which jobs to prioritize.
% TODO: Put numbers here.

### 2015-11-03 pldi-16/evaluation.tex
Implement jiri's comments

\newcommand\mxtest{\texttt{mx\_test}}
\newcommand\tej{\texttt{thr\_exit\_join}}
\newcommand\bct{\texttt{broadcast}}
\newcommand\paraguay{\texttt{paraguay}}
\newcommand\paradise{\texttt{paradise\_lost}}
\newcommand\rwl{\texttt{rwl\_test}}
We tested P2s with 6 multithreaded programs:
% from the 410 test suite % XXX: I would like to say this but this is a lie; figure out what else i can say instead
% each tailored to exercise a different part of the P2 project
\mxtest, for locking algorithm correctness, \tej, a test of thread lifecycle, \bct~and \paraguay, for condvars, \paradise~for semaphores, and \rwl~for r/w locks.
For \mxtest, \paradise, and \paraguay, we used {\tt without\_function} to blacklist {\tt thr\_create}, {\tt thr\_exit}, and {\tt thr\_join},
and for \mxtest~we enabled \landslide's mutex-testing option
(see \sect{\ref{sec:landslide}}).
\newcommand\prisema{\texttt{priority\_sema}}
\newcommand\waitsimple{\texttt{wait\_simple}}
% TODO: Add more test cases
We tested the Pintoses with 2 programs from the class test suite: \prisema, a test of the kernel scheduling algorithm, and \waitsimple, a test of process lifecycle system calls.
For all tests, we used {\tt without\_function} to blacklist PPs on the {\tt malloc} lock.
Finally, all tests were run on 12-core 3.2 GHz Xeon machines with 12GB of RAM.
% XXX: Some pintoses can't run all the tests. So this number is too high.
In total, the test suite comprises 632 unique state spaces, at least 106 of which contain bugs.

@@

{\bf Finer-grained PP subsets.}
% TODO: update these numbers
\quicksand~was able to partially guarantee safety in 97\% of tests with large maximal state spaces.
However, in 2 tests, no more than the minimal state space could be verified,
and in 2 other tests, not even that much.
Larger state spaces often result from finer-grained locking,
which can indicate a more complicated concurrent algorithm requiring more rigorous verification than a program with a single global lock.
%Hence these corner cases are important to consider for future work.
While this work used {\tt within\_function} (\sect{\ref{sec:landslide}}) {\em statically} to restrict where PPs could arise in advance of the test,
we envision future Iterative Deepening implementations could incorporate this method to {\em dynamically} subset PPs further,
making partial verification of such large tests possible.
%enabling partial verification of such large tests. % if need space
%Because our current implementation does not avoid repeated interleavings across state spaces,
%as discussed above,
%we were limited to a small number of very basic PP subsets to statically seed the exploration

>>>>

\newcommand\mxtest{\texttt{mx\_test}}
\newcommand\tej{\texttt{join\_test}}
\newcommand\bct{\texttt{broadcast\_test}}
\newcommand\paraguay{\texttt{signal\_test}}
\newcommand\paradise{\texttt{sem\_test}}
\newcommand\rwl{\texttt{rwl\_test}}
We tested P2s with 6 multithreaded programs:
% from the 410 test suite % XXX: I would like to say this but this is a lie; figure out what else i can say instead
% each tailored to exercise a different part of the P2 project
\mxtest, for locking algorithm correctness, \tej, a test of thread lifecycle, \bct~and \paraguay, for condvars, \paradise~for semaphores, and \rwl~for r/w locks.
For \mxtest, \paradise, and \paraguay, we used {\tt without\_function} to blacklist {\tt thr\_create}, {\tt thr\_exit}, and {\tt thr\_join},
and for \mxtest~we enabled \landslide's mutex-testing option
(see \sect{\ref{sec:landslide}}).
\newcommand\prisema{\texttt{priority\_sema}}
\newcommand\waitsimple{\texttt{wait\_simple}}
% TODO: Add more test cases
We tested the Pintoses with 2 programs from the class test suite: \prisema, a test of the kernel scheduling algorithm, and \waitsimple, a test of process lifecycle system calls.
For all tests, we used {\tt without\_function} to blacklist PPs on the {\tt malloc} lock.
Finally, all tests were run on 12-core 3.2 GHz Xeon machines with 12GB of RAM.
% XXX: Some pintoses can't run all the tests. So this number is too high.
In total, the test suite comprises 632 unique state spaces, at least 106 of which contain bugs.

@@

{\bf Finer-grained PP subsets.}
% TODO: update these numbers
\quicksand~was able to partially guarantee safety in 97\% of tests with large maximal state spaces.
However, in 2 tests, no more than the minimal state space could be verified,
and in 2 other tests, not even that much.
Larger state spaces often result from finer-grained locking,
which can indicate a more intricate concurrent algorithm or an unnecessarily complicated design.
Such programs may require even more rigorous verification than a program with a single global lock.
%Hence these corner cases are important to consider for future work.
While this work used {\tt within\_function} (\sect{\ref{sec:landslide}}) {\em statically} to restrict where PPs could arise in advance of the test,
we envision future Iterative Deepening implementations could incorporate this method to {\em dynamically} subset PPs further,
making partial verification of such large tests possible.
%enabling partial verification of such large tests. % if need space
%Because our current implementation does not avoid repeated interleavings across state spaces,
%as discussed above,
%we were limited to a small number of very basic PP subsets to statically seed the exploration

### 2015-11-03 pldi-16/intro.tex
Implement jiri's comments

% blah blah trite opening sentence
%As parallelism becomes ever more important for achieving high performance in modern-day programs,
%so too do advanced concurrency testing techniques become important for verifying the correctness of those programs.
Concurrency bugs are notoriously hard to find and reproduce because they only appear in specific thread interleavings, which arise at random during normal program execution.
{\em Stateless model checking} \cite{verisoft} offers a method for finding such bugs,
or verifying their absence,
%by systematically executing a program along as many distinct interleavings as possible,
by forcing a program to execute each distinct interleaving,
capturing and controlling this nondeterminism in a finite state space.
Unfortunately, these state spaces explode exponentially in the size of the input program.
Reduction techniques such as Dynamic Partial Order Reduction \cite{dpor} and Maximal Causality Reduction \cite{mcr} expand the limits of feasible test completion,
and search ordering strategies such as Iterative Context Bounding \cite{chess-icb} allow bugs to be found sooner in a given exploration should they exist.

% Can I even make a claim this broad to begin with?
However, all stateless model checkers to date are bound by a fixed set of {\em preemption points}: code locations that define the granularity at which threads interleave.
% TODO: You repeat the following 1.5 sentence in the related work section; maybe you can eliminate this redundancy to save space?
For example, \textsc{CHESS} \cite{chess} preempts only on synchronization operations and library calls, which can miss lock-free shared memory races.
It provides an additional data-race analysis to report any violations of this model;
however, data-race analyses are prone to report false positives
%and benign races
which require annotations or imprecise heuristics to reduce \cite{racerx,tsan,datacollider}.
%
On the other hand, SPIN \cite{spin}
is able to preempt threads around any shared memory access. Such fine granularity would automatically check if each data race is a real bug, but makes full state space completion intractable for even modestly-sized tests.
%
This work shows how to avoid making that tradeoff decision in advance.

@@

This paper's contributions are as follows:
\begin{enumerate}
	\item Iterative Deepening, a new technique for combining data-race analysis with stateless model checking, and \quicksand, an open source implementation;
	\item A new tactic for eliminating one class of false-positive data races, unsound in single-pass analysis but which we prove sound when used with Iterative Deepening;
	%\item Techniques for detecting and flattening cyclic state spaces resulting from ad-hoc while-loop synchronization % TODO: is there actually room for this in the paper?
	\item A large evaluation in which \quicksand~compares favorably to stand-alone data-race detection and stateless model checking approaches.
\end{enumerate}

>>>>

% blah blah trite opening sentence
%As parallelism becomes ever more important for achieving high performance in modern-day programs,
%so too do advanced concurrency testing techniques become important for verifying the correctness of those programs.
Concurrency bugs are notoriously hard to find and reproduce because they only appear in specific thread interleavings, which arise at random during normal program execution.
{\em Stateless model checking} \cite{verisoft} offers a method for finding such bugs,
or verifying their absence,
%by systematically executing a program along as many distinct interleavings as possible,
by forcing a program to execute each distinct interleaving,
capturing and controlling this nondeterminism using a finite state space.
Unfortunately, these state spaces explode exponentially in the size of the input program.
Reduction techniques such as Dynamic Partial Order Reduction \cite{dpor} and Maximal Causality Reduction \cite{mcr} expand the limits of feasible test completion,
and search ordering strategies such as Iterative Context Bounding \cite{chess-icb} allow bugs to be found sooner in a given exploration should they exist.

% Can I even make a claim this broad to begin with?
However, all stateless model checkers to date are bound by a fixed set of {\em preemption points}: code locations that define the granularity at which threads interleave.
% TODO: You repeat the following 1.5 sentence in the related work section; maybe you can eliminate this redundancy to save space?
For example, \textsc{CHESS} \cite{chess} preempts only on synchronization operations and library calls, which can miss lock-free shared memory races.
It provides an additional data-race analysis to report any violations of this model;
however, data-race analyses are prone to report false positives
%and benign races
which require annotations or imprecise heuristics to suppress \cite{racerx,tsan,datacollider}.
%
On the other hand, SPIN \cite{spin}
is able to preempt threads around any shared memory access. Such fine granularity would automatically check if each data race is a real bug, but makes full state space completion intractable for even modestly-sized tests.
%
Configuring a model checker is a tradeoff between schedule coverage and feasibility of completion.
This work shows how to avoid making that tradeoff decision in advance.

@@

This paper's contributions are as follows:
\begin{enumerate}
	\item Iterative Deepening, a new technique for combining data-race analysis with stateless model checking, and \quicksand, an open source implementation of the technique;
	\item A new tactic for eliminating one class of false-positive data races,
		which cannot soundly be used in a single-pass analysis,
		but which we prove correct when used with Iterative Deepening;
		%, unsound in single-pass analysis but which we prove sound when used with Iterative Deepening;
	%\item Techniques for detecting and flattening cyclic state spaces resulting from ad-hoc while-loop synchronization % TODO: is there actually room for this in the paper?
	\item A large evaluation in which \quicksand~compares favorably to stand-alone data-race detection and stateless model checking approaches.
\end{enumerate}
